---
title: "Chapter One Workflow"
author: "Fiona Spooner"
date: "19 October 2017"
output: html_document
---

Step 1 - Calculating Lambdas for each population
```{r, message=FALSE, warning=FALSE}
library(broom)
library(data.table)
library(dplyr)
library(here)
library(lme4)
library(lubridate)
library(mgcv)
library(ncdf4)
library(raster)
library(reshape2)
library(taRifx)
library(zoo)
```

Subsetting the data to exclude fish, populations in marine systems and those without specific locations

```{r, cache=TRUE}

LPI <- read.csv("LPI_pops_20160523_edited.csv")

LPI_pop <- LPI %>% 
  filter(Specific_location == 1 & System !="Marine" & Class != "Actinopterygii"& Class != "Cephalaspidomorphi")

pop_data <- LPI_pop %>% 
  dplyr::select(ID, X1950:X2005)  #selecting out just the ID and the years we are interested in

pop_datab <- (pop_data [,2:57] !="NULL")
points_per_pop1950_2005 <- rowSums(pop_datab)  #calculating the number of years with population estimates
length_id <- data.frame(ID = pop_data$ID,points_per_pop1950_2005)

LPI_pop <- merge(length_id, LPI_pop, by = "ID")

LPI_pop <- LPI_pop %>% 
  filter(points_per_pop1950_2005 >= 2)    #filtering out any population trends with fewer than 2 estimates between 1950 and 2005
  
LPI_pop <- LPI_pop %>% 
  dplyr::select(ID:Binomial, Latitude, Longitude,X1950:X2005)  #selecting out just the ID and the years we are interested in

LPI_pop[LPI_pop == 'NULL'] <- NA

write.csv(LPI_pop, "populations_to_use.csv", row.names = FALSE)

```

Calculating lambdas

```{r}


doFit <- function(sp_name) {
  
  ids <- LPI_pop %>%
    filter(ID == sp_name)

  spid <- ids %>%
    dplyr::select(X1950:X2005)

  colnames(spid) <- 1950:2005              #renaming the date column names as R doesn't like numbered column names
  name <- ids$Binomial
  id <- ids$ID
  points <- ids$points_per_pop1950_2005
  name_id <- paste(name, id, sep = "_") #creating id for naming files of plots
  Date <- as.numeric(colnames(spid))
  spidt <- destring(t(spid))
  #spidt <- as.numeric(as.character(spid))
  time <- length(min(which(!is.na(spidt))):max(which(!is.na(spidt))))
  missing <- time - points

  Year <- Date[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  Population <- spidt[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  Population[Population == 0] <- mean(Population, na.rm = TRUE) * 0.01 #if a population is zero one year thhis is replaced with 1% of the average population estimate - because you can log zeros

  df <- data.frame(Year, Population)

  if (sum(na.omit(df$Population < 1)) > 0) {
    df$Population <- df$Population + 1
  }

  PopN = log10(df$Population)

  if (points >= 6) {
    if (length(na.omit(PopN)) >= 6) {
      SmoothParm = round(length(na.omit(PopN)) / 2)
    } else {
      SmoothParm = 3
    }
  
      mg2 <- mgcv:::gam(PopN ~ s(Year, k = SmoothParm), fx = TRUE)
      pv2 <- predict(mg2, df, type = "response", se = TRUE)
      R_sq2 <- summary(mg2)$r.sq
      model <- 1
      #pv2$fit[pv2$fit <= 0] <- NA
      lambda2 <- diff(pv2$fit)
      lambda_sum2 <- sum(lambda2, na.rm = TRUE)
      lambda_mean2 <- mean(lambda2, na.rm = TRUE)
  
    } else {
    
      SmoothParm <- NA
      PopN = log10(df$Population)
      ml2 <- lm(PopN ~ df$Year)
      R_sq2 <- summary(ml2)$r.sq
      model <- 0
      Pop_interp2 <- na.approx(PopN)
      #Pop_interp2[Pop_interp2 <= 0] <- NA
      lambda2 <- diff(Pop_interp2)
      lambda_sum2 <- sum(lambda2, na.rm = TRUE)
      lambda_mean2 <- mean(lambda2, na.rm = TRUE)
  
  }
  
    res_df = data.frame(
    sp_name = sp_name,
    points = points,
    SmoothParm = SmoothParm,
    r_sq = R_sq2,
    model = model,
    lambda_sum = lambda_sum2,
    lambda_mean = lambda_mean2,
    time = time,
    missing = missing
)

print(res_df)
return(res_df)
}

all_df_list <- lapply(unique(LPI_pop$ID), doFit)

```

```{r, cache=TRUE}
all_df <- do.call(rbind, all_df_list)

write.csv(all_df, "Global_Population_Lambdas.csv", row.names = FALSE)

```

Step 2 - Extracting the Climate data for each population

```{r, cache=TRUE}

CR40s<-brick(here::here("climate_data","cru_ts3.23.1941.1950.tmp.dat.nc"))
CR50s<-brick(here::here("climate_data","cru_ts3.23.1951.1960.tmp.dat.nc"))
CR60s<-brick(here::here("climate_data","cru_ts3.23.1961.1970.tmp.dat.nc"))
CR70s<-brick(here::here("climate_data","cru_ts3.23.1971.1980.tmp.dat.nc"))
CR80s<-brick(here::here("climate_data","cru_ts3.23.1981.1990.tmp.dat.nc"))
CR90s<-brick(here::here("climate_data","cru_ts3.23.1991.2000.tmp.dat.nc"))
CR00s<-brick(here::here("climate_data","cru_ts3.23.2001.2010.tmp.dat.nc"))


```

Subsetting the data to exclude fish, populations in marine systems and those without specific locations
```{r, cache=TRUE}

LPI <- read.csv("populations_to_use.csv")

all_df <- read.csv("Global_Population_Lambdas.csv", stringsAsFactors = FALSE)

LPIsp <- LPI %>% 
  filter(LPI$ID %in% all_df$sp_name)

CR<-stack(CR40s[[97:120]],CR50s,CR60s,CR70s,CR80s,CR90s,CR00s[[1:72]])  #getting years from 1949 - 2006

xy<-data.frame(LPIsp$Longitude, LPIsp$Latitude)

xy<-unique(xy)     #identifying unique locations to extract climate data from 

xy_df<-data.frame(xy)
colnames(xy_df)<-c("lon", "lat")
coordinates(xy_df) <- c("lon", "lat")

```

```{r}
years <- rep(1949:2006, each = 12)
annual_ras <- function(year, rast){
    
  ann_ras <- mean(rast[[which(years == year)]])
  names(ann_ras) <-  paste0("Year_", year)
  return(ann_ras)
}

cr_out <- lapply(X = unique(years), FUN  = annual_ras, rast = CR)

cr_st<- do.call(stack, cr_out)
```

```{r}
mt_out <- raster::extract(cr_st, xy_df, nl = nlayers(cr_st))
mt_df <- data.frame(xy_df@coords, mt_out, row.names = NULL)
melt_df <- melt(mt_df, id = c("lon", "lat"))
colnames(melt_df) <- c("Longitude", "Latitude", "Date", "Mean_T")
write.csv(melt_df, "Global_Mean_Temp.csv", row.names = FALSE)
```

```{r, cache=TRUE}

LPI_pop <- read.csv("populations_to_use.csv")

LPI_pop$ID <- as.integer(LPI_pop$ID)

LPIclim <- vroom::vroom("Global_Mean_Monthly_Temp.csv")
  
LPIclim$Year <- lubridate::year(LPIclim$Date)

LPIclim <- LPIclim[complete.cases(LPIclim),]

LPIclim <-merge(LPI_pop[,c("ID", "Binomial", "Longitude", "Latitude")], LPIclim, by = c("ID","Longitude", "Latitude"))


LPIclim <- LPIclim %>% 
  group_by(ID, Longitude, Latitude, Binomial,Year) %>% 
  summarize(Mean_ann_T = mean(Mean_T))


```

```{r, cache=TRUE}

doMean = function(sp_name) {

  spid2 <- LPI_pop %>% 
    filter(ID == sp_name)   #subsetting the population data by each population
  
  spid <- spid2 %>% 
   dplyr::select(X1950:X2005)
  
  colnames(spid)<-1950:2005              #renaming the date column names as R doesn't like numbered column names
  # nid <- matrix(rep (NA, 10), nrow=1)                
  # colnames(nid) <- 1940:1949
  # spid <- cbind(nid, spid)#renaming the date column names as R doesn't like numbered column names
  climid <- LPIclim %>% 
    filter(ID == sp_name, Year >=  1950 & Year <= 2005) %>% 
    arrange(Year)#subsetting the climate data by each population
  
  name <- spid2$Binomial
  id <- spid2$ID
  points <- spid2$points_per_pop1950_2005
  Date <- as.numeric(colnames(spid))
  spidt <- destring(t(spid))
  mean_change <- NA
  mean_df <- data.frame()
  
  if (any(!is.na(spidt))){
    
  Year <- Date[(min(which(!is.na(spidt)))):max(which(!is.na(spidt)))]
  Population <- spidt[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]

  Mean <- climid$Mean_ann_T[(min(which(!is.na(spidt)))):max(which(!is.na(spidt)))]
  
  mean_change<-sum(diff(Mean))

  if (sum(is.na(Mean))!=length(Mean)){
    
    lm_mean <- lm(Mean~Year)
    #lm_mean<-lm(year_temp$mean_mean~year_temp$Year)
    lm_mean_df <- tidy(lm_mean)[2,]  
    mean_df <- cbind(id,mean_change, lm_mean_df)
    
    } 
  } else{
    
    mean_df <- matrix(c(id,mean_change,NA,NA,NA,NA,NA), nrow=1, ncol=7)
    colnames(mean_df) <- c("id","mean_change" ,"term", "estimate", "std.error", "statistic", "p.value")
    #mean_df<-data.frame(mean_df)
  }
  
  print(sp_name)  
  return(mean_df)
}

all_df_list <- lapply(unique(LPIclim$ID), doMean)


```


```{r, cache=TRUE}
mean_df <- do.call(rbind, all_df_list)
colnames(mean_df) <- c("ID", "Sum_Mean_Change","Term","Estimate","SE","Statistic","p.val")

write.csv(mean_df, "Mean_Temp_Change_LPI_pops.csv", row.names = FALSE)

```


Calculating rates of land use change

```{r, warning=FALSE, message=FALSE}

crop1940<-raster(here::here("hyde_data","crop1940AD.asc"))
gras1940<-raster(here::here("hyde_data","gras1940AD.asc"))

crop1950<-raster(here::here("hyde_data","crop1950AD.asc"))
gras1950<-raster(here::here("hyde_data","gras1950AD.asc"))

crop1960<-raster(here::here("hyde_data","crop1960AD.asc"))
gras1960<-raster(here::here("hyde_data","gras1960AD.asc"))

crop1970<-raster(here::here("hyde_data","crop1970AD.asc"))
gras1970<-raster(here::here("hyde_data","gras1970AD.asc"))

crop1980<-raster(here::here("hyde_data","crop1980AD.asc"))
gras1980<-raster(here::here("hyde_data","gras1980AD.asc"))

crop1990<-raster(here::here("hyde_data","crop1990AD.asc"))
gras1990<-raster(here::here("hyde_data","gras1990AD.asc"))

crop2000<-raster(here::here("hyde_data","crop2000AD.asc"))
gras2000<-raster(here::here("hyde_data","gras2000AD.asc"))

crop2005<-raster(here::here("hyde_data","crop2005AD.asc"))
gras2005<-raster(here::here("hyde_data","gras2005AD.asc"))


rlist<-list(crop1940,crop1950,crop1960,crop1970,crop1980,crop1990,crop2000,crop2005,gras1940,gras1950,gras1960,gras1970,gras1980,gras1990,gras2000,gras2005)
cellareas <- area(rlist[[1]])

for (map in rlist){
  
  map_pcnt <- signif(values(map),digits=4)/cellareas
  map_pcnt_m <- matrix(map_pcnt, nrow=nrow(map), ncol=ncol(map), byrow=TRUE)
  map_pcnt_r <- raster(map_pcnt_m, xmn=map@extent[1], xmx=map@extent[2], ymn=map@extent[3], ymx=map@extent[4])
  writeRaster(map_pcnt_r, here::here("hyde_data", paste0(names(map),"_raster.tif")),overwrite=TRUE)
  print(names(map))
}

crop_s<-stack(list.files(path = here::here("hyde_data"), pattern = "^.*crop*.*.tif$", full.names = TRUE))
gras_s<-stack(list.files(path =  here::here("hyde_data"), pattern = "^.*gras*.*.tif$", full.names = TRUE))

```


```{r, cache=TRUE}
###########LPI data

LPI_LUC <- read.csv("populations_to_use.csv")

latlong <- cbind(LPI_LUC$Longitude,LPI_LUC$Latitude)

x <- latlong[,1]
y <- latlong[,2]
id <- LPI_LUC$ID

xy <- cbind(id,x,y) #creating the grid around the population - 5km either side gives a grid of 121km^2
xy <- na.omit(xy)
id <- xy[,1]

points <- SpatialPoints(cbind(xy[,2], xy[,3]))

xmin <- colFromX(crop_s[[1]], points@coords[,1]) - 1
xmax <- colFromX(crop_s[[1]], points@coords[,1]) + 1
ymin <- rowFromY(crop_s[[1]], points@coords[,2]) - 1
ymax <- rowFromY(crop_s[[1]], points@coords[,2]) + 1    

grid_crop <- data.frame(ID=id,xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax)   #setting extents of grid to extract

grid_crop2 <- merge(grid_crop,LPI_LUC, by="ID")

nrow(grid_crop2)
```


```{r, cache=TRUE, eval=FALSE}

result <- data.frame() #empty result dataframe

grid_crop2[grid_crop2 == 'NULL'] <- NA

for (i in 1:length(grid_crop2$ID)){
  
  id <- grid_crop2[i,1]
  Binomial <- as.character(grid_crop2[i,6])
  spid <- grid_crop2 %>% 
    filter(ID == id) %>% 
    dplyr::select("X1950":"X2005")
  colnames(spid) <- 1950:2005
  
  crop_change <- NA
  gras_change <- NA
  both_change <- NA
  both_sum <- NA
    
  if(any(!is.na(spid))){
    
  Date <- as.numeric(colnames(spid))
  spidt <- destring(t(spid))
  
  Year <- Date[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  Population <- spidt[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]

  crop_check <- crop(crop_s, extent(crop_s, grid_crop2[i,4],grid_crop2[i,5],grid_crop2[i,2],grid_crop2[i,3]))
  gras_check <- crop(gras_s, extent(gras_s, grid_crop2[i,4],grid_crop2[i,5],grid_crop2[i,2],grid_crop2[i,3]))
  
  crop_df <- data.frame(as.matrix(crop_check))
  gras_df <- data.frame(as.matrix(gras_check))
  
  decs <- c("1940","1950","1960","1970","1980","1990","2000","2005")
  colnames(crop_df) <- decs
  colnames(gras_df) <- decs
  
  mean_crop_df <- colMeans(crop_df)
  mean_crop_df2 <- data.frame(mean_crop_df, decs)
  colnames(mean_crop_df2) <- c("mean_crop", "Year")
  
  mean_gras_df <- colMeans(gras_df)
  mean_gras_df2 <- data.frame(mean_gras_df, decs)
  colnames(mean_gras_df2) <- c("mean_grass", "Year")
  
  Year_na <- rep(NA, length(1940:2005))
  Year_all <- 1940:2005
  Yr <- cbind(Year_all, Year_na)
  colnames(Yr) <- c("Year", "Year_NA")
  
  Yr_intrp_crop <- merge(Yr, mean_crop_df2, by="Year", all=T)
  Yr_intrp_gras <- merge(Yr, mean_gras_df2, by="Year", all=T)
  
  min_yr <- min(Year)
  max_yr <- max(Year)
  
  
  if (min_yr != max_yr & min_yr < 2005 & sum(is.na(Yr_intrp_crop$mean_crop))<60) {
    
    Yr_intrp_crop$mean_crop_int <- na.approx(Yr_intrp_crop$mean_crop)
    Yr_intrp_gras$mean_gras_int <- na.approx(Yr_intrp_gras$mean_grass)
    Yr_intrp_crop$both <- Yr_intrp_crop$mean_crop_int + Yr_intrp_gras$mean_gras_int
    
    crop_change <- mean(diff(subset(Yr_intrp_crop, Year >= min_yr & Year <= max_yr)$mean_crop_int))
    gras_change <- mean(diff(subset(Yr_intrp_gras, Year >= min_yr & Year <= max_yr)$mean_gras_int))
    both_change <- mean(diff(subset(Yr_intrp_crop, Year >= min_yr & Year <= max_yr)$both))
    both_sum <- sum(diff(subset(Yr_intrp_crop, Year >= min_yr & Year <= max_yr)$both))
  } 

  }
  
  years <- max(Year)-min(Year)
  final <- cbind(id,Binomial,crop_change, gras_change, both_change, both_sum, years) 
  result <- rbind(final,result)
  print(final)
  
}

write.csv(result, "HYDE_LPI_rate_of_change.csv", row.names = FALSE)

```

```{r, cache=TRUE, warning=FALSE}
lpi <- read.csv("populations_to_use.csv")
bm <- read.csv("Amniote_Database_Aug_2015.csv")
bm$Binomial <- paste(bm$genus, bm$species, sep="_")

lpibm <- merge(lpi, bm, by="Binomial")

lpibm <- data.frame(lpibm$Binomial, lpibm$ID, lpibm$adult_body_mass_g)
colnames(lpibm) <- c("Binomial", "ID", "Body_mass_g")

lpibm$Log_Body_Mass_g <- log10(lpibm$Body_mass_g)

lpibm2 <- lpibm[lpibm$Body_mass_g !=-999,]

lpibm2 <- unique(lpibm2)

lpibm2$Log_Body_Mass_g <- log10(lpibm2$Body_mass_g)

write.csv(lpibm2, "LPI_BodyMass_Amniote_Database.csv", row.names = FALSE)

```

Setting up the modelling

```{r, warning=FALSE, message=FALSE}

#temp <- read.csv("Mean_Temp_Change_LPI_pops.csv", stringsAsFactors = FALSE)
temp <- read.csv("All_LPI_All_Years_Nobuff_1931_moreLPI_end2005.csv")
bm <- read.csv("LPI_BodyMass_Amniote_Database.csv", stringsAsFactors = FALSE)
bm <- bm[,-3]

hyde <- read.csv("HYDE_LPI_rate_of_change.csv", stringsAsFactors = FALSE)
hyde <- hyde %>% 
  dplyr::select(-"Binomial")

temp <- temp %>% 
  dplyr::select("ID", "Estimate")

#temp <- merge(temp, hyde[,c(2,3)], by.x = "ID", by.y = "id")

LPI <- read.csv("LPI_pops_20160523_edited.csv")

pop <- read.csv("Global_Population_Lambdas.csv")

pop <- pop %>% 
  rename(ID = sp_name)

LPI <- LPI %>% 
  dplyr::select("ID","Binomial","Common_name", "Order","Family", "Protected_status", "Country","Region", "System", "Class","Specific_location", "Longitude", "Latitude", "Primary_threat", "Secondary_threat", "Tertiary_threat", "Migratory", "Forest")

df <- merge(merge(temp,bm, by="ID"), merge(LPI, pop, by="ID"),by = c("ID", "Binomial"))

dfd <- merge(df, hyde, by.y = "id", by.x = "ID")

df2 <- dfd %>% 
  filter(!is.na(Estimate) & r_sq >= 0.4999999  & time >= 5 & System!="Marine" 
        & Specific_location == 1 & !is.na(both_change) & !is.na(Log_Body_Mass_g) & Class=="Aves")

nrow(df2)

table(df2$Protected_status)
df2$Protected_status[df2$Protected_status == "No (large survey area)"]<-"No"
df2$Protected_status[df2$Protected_status == "No (area surrounding PA)"]<-"No"

```

```{r}

sp_dups <- df2 %>% 
  group_by(Longitude, Latitude) %>% 
  summarise(n = n()) 

sp_dups$loc_id <- 1:length(sp_dups$Longitude)

sp_dups_df<-merge(sp_dups, df2, by=c("Longitude","Latitude"))

parm_df <- sp_dups_df %>% 
  dplyr::select("ID","Estimate", "both_change", "Log_Body_Mass_g")

parm_scale <- scale(parm_df[,c("Estimate", "both_change", "Log_Body_Mass_g")])       #use the scaling factors at the bottom of these to scale the rasters

parm_id <- parm_df[,"ID"]

parm_df_scale <- data.frame(parm_id,parm_scale)

colnames(parm_df_scale) <- c("ID","mean_slope_scale", "change_rate_scale", "Bodymass_scale")

dt <- merge(sp_dups_df, parm_df_scale, by="ID")

nrow(dt)
```


```{r, cache=TRUE, warning=FALSE, message=FALSE}
source("rsquaredglmm.R")

  
  m0<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+Bodymass_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m0f<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+Bodymass_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)

  m0a<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+Bodymass_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m0af<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+Bodymass_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  
  m0b<-lmer(lambda_mean ~ change_rate_scale+Bodymass_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m0bf<-lmer(lambda_mean ~ change_rate_scale+Bodymass_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  
  m0c<-lmer(lambda_mean ~ mean_slope_scale+Bodymass_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m0cf<-lmer(lambda_mean ~ mean_slope_scale+Bodymass_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  
  m0d<-lmer(lambda_mean ~ Bodymass_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m0df<-lmer(lambda_mean ~ Bodymass_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  
  m1<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m1f<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)

  m1a<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m1af<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)

  m1b<-lmer(lambda_mean ~ change_rate_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m1bf<-lmer(lambda_mean ~ change_rate_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)  

  m1c<-lmer(lambda_mean ~ mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m1cf<-lmer(lambda_mean ~ mean_slope_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  mnull<-lmer(lambda_mean ~ 1+(1|Binomial)+(1|loc_id),data=dt, REML=F)
```
  
Model Comparison
```{r, warning=FALSE, message=FALSE}
  # #Weights
  library(MuMIn)
  library(lme4)  

  #msAICc <- model.sel(m1,m1a,m1b,m1c,mnull)
  #msAICc <- model.sel(m0,m0a,m0b,m0c,m0d,m1,m1a,m1b,m1c,m0f,m0af,m0bf,m0cf,m0df,m1f,m1af,m1bf,m1cf,mnull)  
  msAICc <- model.sel(m0,m0a,m0b,m0c,m0d,m1,m1a,m1b,m1c,mnull)

  msAICc$model<-rownames(msAICc)
  msAICc<-data.frame(msAICc)
  msAICc
  
  AIC(m0,m0a,m0b,m0c,m1,m1a,m1b,m1c,mnull,m0f,m0af,m0bf,m0cf,m0df,m1f,m1af,m1bf,m1cf)
  
  #Rsq
  #models_list<-list(m0,m0a,m0b,m0c,m0d,m1,m1a,m1b,m1c,mnull,m0f,m0af,m0bf,m0cf,m0df,m1f,m1af,m1bf,m1cf)
  models_list<-list(m0,m0a,m0b,m0c,m0d,m1,m1a,m1b,m1c,mnull)
  modelsR<-lapply(models_list,rsquared.glmm)
  modelsRsq <- matrix(unlist(modelsR), ncol=6, byrow=T)
  rownames(modelsRsq)<-c("m0","m0a","m0b","m0c","m0d","m1","m1a","m1b","m1c","mnull")

  modelsRsq
    
  library(MuMIn)
  var_imp<-summary(model.avg(models_list))

  #mav<-model.avg(models_list, subset = cumsum(weight) <= 0.95)
  mav<-model.avg(models_list, subset = delta <= 6)
  
  smav<-summary(mav)
  
  coef_av<-smav$coefmat.subset[,"Estimate"]
  coef_df<-data.frame(coef_av)
  coef_df$lowCI<-confint(mav)[,1]
  coef_df$highCI<-confint(mav)[,2]
  coef_df
  
  
  coef_pcnt<-data.frame(((10^coef_df) - 1)*100)
  coef_pcnt
  

  coef_pcnt$Var_name<-rownames(coef_pcnt)
  # library(plotrix)
  # 
  # plotCI(1:5, y=coef_pcnt$coef_av, ui=coef_pcnt$highCI, li=coef_pcnt$lowCI, ylab="Annual Population Change (%)", xlab="" ,xaxt = "n", 
  #        main="Birds", lwd=1, ylim=c(min(coef_pcnt$lowCI*1.1), max(coef_pcnt$highCI*1.2)))
  # axis(1, at=1:5, labels=rownames(coef_pcnt), las=2)
  # abline(h=0, col="red", lty =2)

  #AIC
  AICs<-c(AIC(m0),AIC(m0a),AIC(m0b),AIC(m0c),AIC(m0d),AIC(m1),AIC(m1a),AIC(m1b),AIC(m1c),AIC(mnull))
  mnames<-c("LUC*MTC+BM","LUC+MTC+BM", "LUC+BM", "MTC+BM","BM","LUC*MTC","LUC+MTC", "LUC", "MTC", "NULL")
  
  AIC_diff<-AICs - AIC(mnull)
  del_AIC_df<-data.frame(AIC_diff, mnames)
  del_AIC_df<-del_AIC_df[order(del_AIC_df$AIC_diff),]
  del_AIC_df
  
  
library(coefplot)
library(ggplot2)

coef_pcnt$val<-1:5
coef_pcnt$var_name <- factor(coef_pcnt$Var_name, levels = coef_pcnt$Var_name[order(coef_pcnt$val)])
  
  p1<-ggplot(coef_pcnt)
  p1<- p1 + geom_hline(yintercept = 0, colour=gray(1/2), lty=2)
  p1<- p1 + geom_linerange(aes(x=Var_name, ymin=lowCI, ymax=highCI), lwd=1.5, position = position_dodge(width=1/2))
  p1<- p1 + geom_pointrange(aes(x= Var_name, y=coef_av, ymin=lowCI, ymax=highCI), lwd=1, position=position_dodge(width=1/2), shape=21, fill="White")
  p1<- p1 + scale_y_continuous(breaks=seq(-8, 4, 2)) +theme_bw() + labs(y = "Annual Population Change (%)", x = "Variable") + theme(legend.title=element_blank(), text = element_text(size=20),axis.title.x = element_text(margin = unit(c(5, 5, 0, 0), "mm")))
  p1<-p1 + coord_flip()
  print(p1)
  
  
#coef_pcnt$Class<-"Birds"  
#coef_pcnt$Class<-"Mammals"  

#coef_pcntb<-coef_pcnt
#coef_pcntm<-coef_pcnt  


```

```{r, eval=FALSE, echo=FALSE}
coef_both<-rbind(coef_pcntb[,c(1,2,3,4,7)], coef_pcntm[,c(1,2,3,4,7)])

coef_both$Var_name
coef_both$Var_name[coef_both$Var_name == "(Intercept)"] <- "Intercept"
coef_both$Var_name[coef_both$Var_name == "mean_slope_scale"] <- "MTC"
coef_both$Var_name[coef_both$Var_name == "change_rate_scale"] <- "LUC"
coef_both$Var_name[coef_both$Var_name == "change_rate_scale:mean_slope_scale"] <- "MTC*LUC"
coef_both$Var_name[coef_both$Var_name == "Bodymass_scale"] <- "BM"

#write.csv(coef_both, "Model_Average_coefs4.csv")
coefs_both<-read.csv("Model_Average_coefs4.csv")

library(ggplot2)
p1<-ggplot(coefs_both, aes(colour=Class))
p1<- p1 + geom_hline(yintercept = 0, colour=gray(1/2), lty=2)
p1<- p1 + geom_linerange(aes(x=Var_name, ymin=lowCI, ymax=highCI), lwd=2.5, position = position_dodge(width=2/3))
p1<- p1 + geom_pointrange(aes(x= Var_name, y=coef_av, ymin=lowCI, ymax=highCI), lwd=2, position=position_dodge(width=2/3), shape=21, fill="White")
p1<- p1 + scale_y_continuous(breaks=seq(-8, 6, 2), limits=(c(-9,5))) +theme_bw() + labs(y = "Annual Population Rate of Population Change (%)", x = "") + theme(legend.position="none",text=element_text(size=20),axis.text.x=element_text(size=20) , axis.title.x = element_text(margin = unit(c(5, 0, 0, 0), "mm")))
p1<- p1 + scale_color_manual(values=c("black", "black"))

print(p1)

```