---
title: "Chapter One Workflow"
author: "Fiona Spooner"
date: "19 October 2017"
output: html_document
---

Step 1 - Calculating Lambdas for each population
```{r, message=FALSE, warning=FALSE}
devtools::install_github("timnewbold/GISOperations")

library(broom)
library(dplyr)
library(devtools)
library(GISOperations)
library(here)
library(lubridate)
library(mgcv)
library(raster)
library(reshape2)
library(taRifx)
library(zoo)
```

Subsetting the data to exclude fish, populations in marine systems and those without specific locations

```{r, cache=TRUE}

LPI <- read.csv("LPI_pops_20160523_edited.csv")

LPI_pop <- LPI %>% 
  filter(Specific_location==1 & System !="Marine" & Class != "Actinopterygii"& Class != "Cephalaspidomorphi")

pop_data <- LPI_pop %>% 
  dplyr::select(ID, X1950:X2005)  #selecting out just the ID and the years we are interested in

pop_datab <- (pop_data [,2:57] !="NULL")
points_per_pop1950_2005 <- rowSums(pop_datab)  #calculating the number of years with population estimates
length_id <- data.frame(ID = pop_data$ID,points_per_pop1950_2005)

LPI_pop <- merge(length_id, LPI_pop, by = "ID")

LPI_pop <- LPI_pop %>% 
  filter(points_per_pop1950_2005 >= 2)    #filtering out any population trends with fewer than 2 estimates between 1950 and 2005
  
LPI_pop <- LPI_pop %>% 
  dplyr::select(ID:Binomial, X1950:X2005)  #selecting out just the ID and the years we are interested in

LPI_pop[LPI_pop == 'NULL'] <- NA

#write.csv(LPI_pop, "populations_to_use.csv", row.names = FALSE)

```

Calculating lambdas

```{r}

doFit <- function(sp_name) {
 
  ids <- LPI_pop %>% 
    filter(ID == sp_name)

  spid <- ids %>% 
   dplyr::select(X1950:X2005)
  
  colnames(spid)<-1950:2005              #renaming the date column names as R doesn't like numbered column names
  name <- ids$Binomial
  id <- ids$ID
  points <- ids$points_per_pop1950_2005
  name_id <- paste(name, id, sep="_") #creating id for naming files of plots
  Date <- as.numeric(colnames(spid))
  spidt <- destring(t(spid))
  time <- length(min(which(!is.na(spidt))):max(which(!is.na(spidt))))
  missing <- time - points
  
  Year<-Date[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  Population<-spidt[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  Population[Population == 0] <- mean(Population, na.rm=TRUE)*0.01 #if a population is zero one year thhis is replaced with 1% of the average population estimate - because you can log zeros
  
  df<-data.frame(Year,Population)
  
  if (sum(na.omit(df$Population < 1))>0) {
    df$Population<-df$Population + 1
  } 
    
  
  if (points >=6) {           
    PopN = log10(df$Population)
    if (length(na.omit(PopN)) >=6) {
      SmoothParm = round(length(na.omit(PopN))/2)    
    } else {
      SmoothParm=3
    }

    mg2<-mgcv:::gam(PopN ~ s(Year, k=SmoothParm), fx=TRUE)
    pv2 <- predict(mg2,df,type="response",se=TRUE) 
    R_sq2<-summary(mg2)$r.sq
    model<-1
    pv2$fit[pv2$fit <= 0] <- NA
    lambda2<-diff(pv2$fit)
    lambda_sum2<-sum(lambda2, na.rm=TRUE)
    lambda_mean2<-mean(lambda2, na.rm=TRUE)

  } else {
    SmoothParm<-NA
    PopN = log10(df$Population)
    ml2<-lm(PopN~df$Year)
    R_sq2<-summary(ml2)$r.sq
    model<-0
    Pop_interp2<-na.approx(PopN)
    Pop_interp2[Pop_interp2<=0] <- NA
    lambda2<-diff(Pop_interp2)
    lambda_sum2<-sum(lambda2, na.rm=TRUE)
    lambda_mean2<-mean(lambda2, na.rm=TRUE)
 
  }

res_df = data.frame(sp_name=sp_name, points=points, SmoothParm=SmoothParm, r_sq=R_sq2, model=model,lambda_sum=lambda_sum2,lambda_mean=lambda_mean2,time=time, missing=missing)

#print(res_df)
return(res_df)
}

all_df_list <- lapply(unique(LPI_pop$ID), doFit)

```

```{r, cache=TRUE}
all_df <- do.call(rbind, all_df_list)

write.csv(all_df, "Global_Population_Lambdas.csv", row.names = FALSE)

```

Step 2 - Extracting the Climate data for each population

```{r, cache=TRUE}
CR30s<-brick(here::here("climate_data","cru_ts3.23.1931.1940.tmp.dat.nc"))
CR40s<-brick(here::here("climate_data","cru_ts3.23.1941.1950.tmp.dat.nc"))
CR50s<-brick(here::here("climate_data","cru_ts3.23.1951.1960.tmp.dat.nc"))
CR60s<-brick(here::here("climate_data","cru_ts3.23.1961.1970.tmp.dat.nc"))
CR70s<-brick(here::here("climate_data","cru_ts3.23.1971.1980.tmp.dat.nc"))
CR80s<-brick(here::here("climate_data","cru_ts3.23.1981.1990.tmp.dat.nc"))
CR90s<-brick(here::here("climate_data","cru_ts3.23.1991.2000.tmp.dat.nc"))
CR00s<-brick(here::here("climate_data","cru_ts3.23.2001.2010.tmp.dat.nc"))
```

Subsetting the data to exclude fish, populations in marine systems and those without specific locations
```{r, cache=TRUE}

LPI <- read.csv("LPI_pops_20160523_edited.csv")

all_df <- read.csv("Global_Population_Lambdas.csv", stringsAsFactors = FALSE)

LPIsp <- LPI %>% 
  filter(LPI$ID %in% all_df$sp_name)

CR<-stack(CR30s,CR40s,CR50s,CR60s,CR70s,CR80s,CR90s,CR00s)

xy<-data.frame(LPIsp$Longitude, LPIsp$Latitude)

xy<-unique(xy)     #identifying unique locations to extract climate data from 

xy_df<-data.frame(xy)
colnames(xy_df)<-c("lon", "lat")
coordinates(xy_df) <- c("lon", "lat")

```

```{r}

years <- rep(1931:2010, each = 12)

annual_ras <- function(year, rast){
    
  ann_ras <- mean(rast[[which(years == year)]])
  names(ann_ras) <-  paste0("Year_", year)
  return(ann_ras)

  }

cr_out <- lapply(X = unique(years), FUN  = annual_ras, rast = CR)

cr_st<- do.call(stack, cr_out)

```

```{r}

mt_out <- raster::extract(cr_st, xy_df, nl = nlayers(cr_st))

mt_df <- data.frame(xy_df@coords, mt_out, row.names = NULL)

melt_df <- melt(mt_df, id = c("lon", "lat"))

colnames(melt_df) <- c("Longitude", "Latitude", "Date", "Mean_T")

#write.csv(melt_df, "Global_Mean_Temp.csv", row.names = FALSE)

```


```{r, cache=TRUE}
LPI_pop <- read.csv("populations_to_use.csv")

LPI_pop$ID <- as.integer(LPI_pop$ID)

LPIclim <- read.csv("Global_Mean_Temp.csv", stringsAsFactors = FALSE)
  
LPIclim$Year <- as.numeric(gsub("Year_","",LPIclim$Date))

LPIclim <- LPIclim[complete.cases(LPIclim),]

LPIclim <-merge(LPI_pop[,c("ID", "Binomial", "Longitude", "Latitude")], LPIclim, by = c("Longitude", "Latitude"))



```

```{r, cache=TRUE}

doMean = function(sp_name) {

  spid2 <- LPI_pop %>% 
    filter(ID == sp_name)   #subsetting the population data by each population
  
  spid <- spid2 %>% 
   dplyr::select(X1950:X2005)
  
  colnames(spid)<-1950:2005              #renaming the date column names as R doesn't like numbered column names
  nid <- matrix(rep (NA, 10), nrow=1)                
  colnames(nid) <- 1940:1949
  spid <- cbind(nid, spid)#renaming the date column names as R doesn't like numbered column names
  climid <- LPIclim %>% 
    filter(ID == sp_name, Year >=  1940 & Year <= 2005) %>% 
    arrange(Year)#subsetting the climate data by each population
  
  name <- spid2$Binomial
  id <- spid2$ID
  points <- spid2$points_per_pop1950_2005
  Date <- as.numeric(colnames(spid))
  spidt <- destring(t(spid))
  mean_change <- NA
  
  if (any(!is.na(spidt))){
    
  Year <- Date[(min(which(!is.na(spidt)))):max(which(!is.na(spidt)))]
  Population <- spidt[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]

  Mean <- climid$Mean_T[(min(which(!is.na(spidt)))):max(which(!is.na(spidt)))]
  
  mean_change<-sum(diff(Mean))

  if (sum(is.na(Mean))!=length(Mean)){
    
    lm_mean <- lm(Mean~Year)
    #lm_mean<-lm(year_temp$mean_mean~year_temp$Year)
    lm_mean_df <- tidy(lm_mean)[2,]  
    mean_df <- cbind(id,mean_change, lm_mean_df)
    
    } 
  } else{
    
    mean_df <- matrix(c(id,mean_change,NA,NA,NA,NA,NA), nrow=1, ncol=7)
    colnames(mean_df) <- c("id","mean_change" ,"term", "estimate", "std.error", "statistic", "p.value")
    #mean_df<-data.frame(mean_df)
  }
  
  print(sp_name)  
  return(mean_df)
}

all_df_list <- lapply(unique(LPI_pop$ID), doMean)


```


```{r, cache=TRUE}
mean_df <- do.call(rbind, all_df_list)
colnames(mean_df) <- c("ID", "Sum_Mean_Change","Term","Estimate","SE","Statistic","p.val")

#write.csv(mean_df, "Mean_Temp_Change_LPI_pops.csv", row.names = FALSE)

```


Calculating rates of land use change

```{r, warning=FALSE, message=FALSE}

crop1940<-raster(here::here("hyde_data","crop1940AD.asc"))
gras1940<-raster(here::here("hyde_data","gras1940AD.asc"))

crop1950<-raster(here::here("hyde_data","crop1950AD.asc"))
gras1950<-raster(here::here("hyde_data","gras1950AD.asc"))

crop1960<-raster(here::here("hyde_data","crop1960AD.asc"))
gras1960<-raster(here::here("hyde_data","gras1960AD.asc"))

crop1970<-raster(here::here("hyde_data","crop1970AD.asc"))
gras1970<-raster(here::here("hyde_data","gras1970AD.asc"))

crop1980<-raster(here::here("hyde_data","crop1980AD.asc"))
gras1980<-raster(here::here("hyde_data","gras1980AD.asc"))

crop1990<-raster(here::here("hyde_data","crop1990AD.asc"))
gras1990<-raster(here::here("hyde_data","gras1990AD.asc"))

crop2000<-raster(here::here("hyde_data","crop2000AD.asc"))
gras2000<-raster(here::here("hyde_data","gras2000AD.asc"))

crop2005<-raster(here::here("hyde_data","crop2005AD.asc"))
gras2005<-raster(here::here("hyde_data","gras2005AD.asc"))


rlist<-list(crop1940,crop1950,crop1960,crop1970,crop1980,crop1990,crop2000,crop2005,gras1940,gras1950,gras1960,gras1970,gras1980,gras1990,gras2000,gras2005)
cellareas <- DegreeCellAreaKM(lat=coordinates(crop1940)[,2],height=res(crop1940)[2],width=res(crop1940)[1])

for (map in rlist){
  
  map_pcnt<-signif(values(map),digits=4)/cellareas
  map_pcnt_m<-matrix(map_pcnt, nrow=nrow(map), ncol=ncol(map), byrow=TRUE)
  map_pcnt_r<-raster(map_pcnt_m, xmn=map@extent[1], xmx=map@extent[2], ymn=map@extent[3], ymx=map@extent[4])
  writeRaster(map_pcnt_r, here::here("hyde_data", paste0(names(map),"_raster.tif")),overwrite=TRUE)
  print(names(map))
}

crop_s<-stack(list.files(path = here::here("hyde_data"), pattern = "^.*crop*.*.tif$", full.names = TRUE))
gras_s<-stack(list.files(path =  here::here("hyde_data"), pattern = "^.*gras*.*.tif$", full.names = TRUE))

```


```{r, cache=TRUE}
###########LPI data

LPI_LUC<-read.csv("populations_to_use.csv")

latlong<-cbind(LPI_LUC$Longitude,LPI_LUC$Latitude)

x<-latlong[,1]
y<-latlong[,2]

xy<-cbind(id,x,y) #creating the grid around the population - 5km either side gives a grid of 121km^2
xy<-na.omit(xy)
id<-xy[,1]

points<-SpatialPoints(cbind(xy[,2], xy[,3]))


xmin<- colFromX(crop_s[[1]], points[1:length(points),]) - 1
xmax<- colFromX(crop_s[[1]], points[1:length(points),]) + 1
ymin<- rowFromY(crop_s[[1]], points[1:length(points),]) - 1
ymax<- rowFromY(crop_s[[1]], points[1:length(points),]) + 1    


grid_crop<-data.frame(ID=id,xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax)   #setting extents of grid to extract

grid_crop2<-merge(grid_crop,LPI_LUC, by="ID")

grid_crop2<-subset(grid_crop2, Specific_location ==1) 

nrow(grid_crop2)
```


```{r, cache=TRUE, eval=FALSE}
library(taRifx)
library(zoo)

result <- data.frame() #empty result dataframe

for (i in 1:length(grid_crop2$ID)){
  
  ID<-grid_crop2[i,1]
  Binomial<-as.character(grid_crop2[i,6])
  spid = grid_crop2[i,69:134]                     #subsetting only the dates
  colnames(spid)<-1950:2015
  
  Date<-as.numeric(colnames(spid))
  spidt<-destring(t(spid))
  
  Year<-Date[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]
  Population<-spidt[min(which(!is.na(spidt))):max(which(!is.na(spidt)))]

  crop_check<-crop(crop_s, extent(crop_s, grid_crop2[i,4],grid_crop2[i,5],grid_crop2[i,2],grid_crop2[i,3]))
  gras_check<-crop(gras_s, extent(gras_s, grid_crop2[i,4],grid_crop2[i,5],grid_crop2[i,2],grid_crop2[i,3]))
  
  crop_df<-data.frame(as.matrix(crop_check))
  gras_df<-data.frame(as.matrix(gras_check))
  
  decs<-c("1940","1950","1960","1970","1980","1990","2000","2005")
  colnames(crop_df)<-decs
  colnames(gras_df)<-decs
  
  mean_crop_df<-colMeans(crop_df)
  mean_crop_df2<-data.frame(mean_crop_df, decs)
  colnames(mean_crop_df2)<-c("mean_crop", "Year")
  
  mean_gras_df<-colMeans(gras_df)
  mean_gras_df2<-data.frame(mean_gras_df, decs)
  colnames(mean_gras_df2)<-c("mean_grass", "Year")
  
  Year_na<-rep(NA, length(1940:2005))
  Year_all<-1940:2005
  Yr<-cbind(Year_all, Year_na)
  colnames(Yr)<-c("Year", "Year_NA")
  
  Yr_intrp_crop<-merge(Yr, mean_crop_df2, by="Year", all=T)
  Yr_intrp_gras<-merge(Yr, mean_gras_df2, by="Year", all=T)
  
  min_yr<-min(Year)
  max_yr<-max(Year)
  
  
  if (min_yr != max_yr & min_yr < 2005 & sum(is.na(Yr_intrp_crop$mean_crop))<60) {
    
    Yr_intrp_crop$mean_crop_int<-na.approx(Yr_intrp_crop$mean_crop)
    Yr_intrp_gras$mean_gras_int<-na.approx(Yr_intrp_gras$mean_grass)
    Yr_intrp_crop$both<-Yr_intrp_crop$mean_crop_int + Yr_intrp_gras$mean_gras_int
    
    crop_change<-mean(diff(subset(Yr_intrp_crop, Year >= min_yr & Year <= max_yr)$mean_crop_int))
    gras_change<-mean(diff(subset(Yr_intrp_gras, Year >= min_yr & Year <= max_yr)$mean_gras_int))
    both_change<-mean(diff(subset(Yr_intrp_crop, Year >= min_yr & Year <= max_yr)$both))
    both_sum<-sum(diff(subset(Yr_intrp_crop, Year >= min_yr & Year <= max_yr)$both))
  }else{
    crop_change<-NA
    gras_change<-NA
    both_change<-NA
    both_sum<-NA
  }
  
  years<-max(Year)-min(Year)
  final<-cbind(ID,Binomial,crop_change, gras_change, both_change, both_sum, years) 
  result<-rbind(final,result)
  print(final)
  
}

#write.csv(result, "Hyde_crop_pasture_annual_change_sum.csv")

```

```{r, cache=TRUE, warning=FALSE}
lpi<-read.csv("LPI_pops_20160523_edited.csv")
bm<-read.csv("Amniote_Database_Aug_2015.csv")
bm$Binomial<-paste(bm$genus, bm$species, sep="_")

lpibm<-merge(lpi, bm, by="Binomial")

lpibm<-data.frame(lpibm$Binomial, lpibm$ID, lpibm$adult_body_mass_g)
head(lpibm)
colnames(lpibm)<-c("Binomial", "ID", "Body_mass_g")

lpibm$Log_Body_Mass_g<-log10(lpibm$Body_mass_g)

lpibm2<-lpibm[lpibm$Body_mass_g !=-999,]


lpibm2<-unique(lpibm2)

lpibm2$Log_Body_Mass_g<-log10(lpibm2$Body_mass_g)

#write.csv(lpibm2, "LPI_BodyMass_Amniote_Database.csv")

```

Setting up the modelling

```{r, cache=TRUE, warning=FALSE, message=FALSE}

temp<-read.csv("All_LPI_All_Years_Nobuff_1931_moreLPI_end2005.csv")

body4<-read.csv("LPI_BodyMass_Amniote_Database.csv")

body4<-body4[,-3]
hyde<-read.csv("Hyde_crop_pasture_annual_change.csv")

hyde2<-read.csv("Hyde_crop_pasture_annual_change_sum.csv")

temp<-temp[,c("ID", "Estimate")]
temp<-merge(temp, hyde[,c(2,3)], by="ID")

LPI<-read.csv("LPI_pops_20160523_edited.csv")

pop<-read.csv("Global_Population_Trends_Rsq_Lambda_07_10_16.csv")



LPI<-LPI[,c("ID","Binomial","Common_name", "Order","Family", "Protected_status", "Country","Region", "System", "Class","Specific_location", "Longitude", "Latitude", "Primary_threat", "Secondary_threat", "Tertiary_threat", "Migratory", "Forest")]

df<-merge(merge(temp,body4[,c(2:4)], by="Binomial"), merge(LPI, pop, by="ID"),by="ID")


dfd<-merge(df, hyde[,c(-1,-3)], by="ID")

head(dfd)

nrow(df)
nrow(dfd) 
dfu<-unique(dfd)

df2<-subset(dfu, !is.na(Estimate) & r_sq >= 0.4999999  &length_time >=5& System!="Marine" 
            &Specific_location == 1 &!is.na(both_change) & !is.na(Log_Body_Mass_g)& Class=="Aves")
nrow(df2)

table(df2$Protected_status)
df2$Protected_status[df2$Protected_status == "No (large survey area)"]<-"No"
df2$Protected_status[df2$Protected_status == "No (area surrounding PA)"]<-"No"

colnames(df2)[2]<-"Binomial"

#df2[is.na(df2$lambda_mean),]$lambda_mean<-0

nrow(df2)

#write.csv(df2, "mammal_data_for_prediction.csv")

library(plyr)
sp_dups<-data.frame(ddply(df2,.(Longitude,Latitude),nrow))
sp_dups$loc_id<-1:length(sp_dups$Longitude)
sp_dups_df<-merge(sp_dups, df2, by=c("Longitude","Latitude"))

library(data.table)
dt = as.data.table(sp_dups_df)

parm_df<-sp_dups_df[,c("ID","Estimate", "both_change", "Log_Body_Mass_g")]  ##ID, land use, and climate  use "LUC_dist" or "Nat_change" for purely annual change in summed primary, secondary and other

parm_mat<-as.matrix(parm_df)
parm_scale<-scale(parm_mat[,c("Estimate", "both_change", "Log_Body_Mass_g")])       #use the scaling factors at the bottom of these to scale the rasters

parm_id<-parm_mat[,"ID"]

parm_df_scale<-data.frame(parm_id,parm_scale)

colnames(parm_df_scale)<-c("ID","mean_slope_scale", "change_rate_scale", "Bodymass_scale")

sp_df_scale<-merge(sp_dups_df, parm_df_scale, by="ID")

dt<-data.table(sp_df_scale)

length(unique(dt$loc_id))

nrow(dt)

#write.csv(dt, "Mammals_scaled_ready_for_models.csv")
```


```{r, cache=TRUE, warning=FALSE, message=FALSE}
source("rsquaredglmm.R")

  library(lme4) 
  
  m0<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+Bodymass_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m0f<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+Bodymass_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)

  m0a<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+Bodymass_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m0af<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+Bodymass_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  
  m0b<-lmer(lambda_mean ~ change_rate_scale+Bodymass_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m0bf<-lmer(lambda_mean ~ change_rate_scale+Bodymass_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  
  m0c<-lmer(lambda_mean ~ mean_slope_scale+Bodymass_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m0cf<-lmer(lambda_mean ~ mean_slope_scale+Bodymass_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  
  m0d<-lmer(lambda_mean ~ Bodymass_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m0df<-lmer(lambda_mean ~ Bodymass_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  
  m1<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m1f<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+change_rate_scale:mean_slope_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)

  m1a<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m1af<-lmer(lambda_mean ~ change_rate_scale+mean_slope_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)

  m1b<-lmer(lambda_mean ~ change_rate_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m1bf<-lmer(lambda_mean ~ change_rate_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)  

  m1c<-lmer(lambda_mean ~ mean_slope_scale+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  m1cf<-lmer(lambda_mean ~ mean_slope_scale+Protected_status+(1|Binomial)+(1|loc_id),data=dt, REML=F)
  mnull<-lmer(lambda_mean ~ 1+(1|Binomial)+(1|loc_id),data=dt, REML=F)
```
  
Model Comparison
```{r, warning=FALSE, message=FALSE}
  # #Weights
  library(MuMIn)
  library(lme4)  

  #msAICc <- model.sel(m1,m1a,m1b,m1c,mnull)
  #msAICc <- model.sel(m0,m0a,m0b,m0c,m0d,m1,m1a,m1b,m1c,m0f,m0af,m0bf,m0cf,m0df,m1f,m1af,m1bf,m1cf,mnull)  
  msAICc <- model.sel(m0,m0a,m0b,m0c,m0d,m1,m1a,m1b,m1c,mnull)

  msAICc$model<-rownames(msAICc)
  msAICc<-data.frame(msAICc)
  msAICc
  
  AIC(m0,m0a,m0b,m0c,m1,m1a,m1b,m1c,mnull,m0f,m0af,m0bf,m0cf,m0df,m1f,m1af,m1bf,m1cf)
  
  #Rsq
  #models_list<-list(m0,m0a,m0b,m0c,m0d,m1,m1a,m1b,m1c,mnull,m0f,m0af,m0bf,m0cf,m0df,m1f,m1af,m1bf,m1cf)
  models_list<-list(m0,m0a,m0b,m0c,m0d,m1,m1a,m1b,m1c,mnull)
  modelsR<-lapply(models_list,rsquared.glmm)
  modelsRsq <- matrix(unlist(modelsR), ncol=6, byrow=T)
  rownames(modelsRsq)<-c("m0","m0a","m0b","m0c","m0d","m1","m1a","m1b","m1c","mnull")

  modelsRsq
    
  library(MuMIn)
  var_imp<-summary(model.avg(models_list))

  #mav<-model.avg(models_list, subset = cumsum(weight) <= 0.95)
  mav<-model.avg(models_list, subset = delta <= 6)
  
  smav<-summary(mav)
  
  coef_av<-smav$coefmat.subset[,"Estimate"]
  coef_df<-data.frame(coef_av)
  coef_df$lowCI<-confint(mav)[,1]
  coef_df$highCI<-confint(mav)[,2]
  coef_df
  
  
  coef_pcnt<-data.frame(((10^coef_df) - 1)*100)
  coef_pcnt
  

  coef_pcnt$Var_name<-rownames(coef_pcnt)
  library(plotrix)
  
  plotCI(1:5, y=coef_pcnt$coef_av, ui=coef_pcnt$highCI, li=coef_pcnt$lowCI, ylab="Annual Population Change (%)", xlab="" ,xaxt = "n", 
         main="Birds", lwd=1, ylim=c(min(coef_pcnt$lowCI*1.1), max(coef_pcnt$highCI*1.2)))
  axis(1, at=1:5, labels=rownames(coef_pcnt), las=2)
  abline(h=0, col="red", lty =2)

  #AIC
  AICs<-c(AIC(m0),AIC(m0a),AIC(m0b),AIC(m0c),AIC(m0d),AIC(m1),AIC(m1a),AIC(m1b),AIC(m1c),AIC(mnull))
  mnames<-c("LUC*MTC+BM","LUC+MTC+BM", "LUC+BM", "MTC+BM","BM","LUC*MTC","LUC+MTC", "LUC", "MTC", "NULL")
  
  AIC_diff<-AICs - AIC(mnull)
  del_AIC_df<-data.frame(AIC_diff, mnames)
  del_AIC_df<-del_AIC_df[order(del_AIC_df$AIC_diff),]
  del_AIC_df
  
  
library(coefplot)
library(ggplot2)

coef_pcnt$val<-1:5
coef_pcnt$var_name <- factor(coef_pcnt$Var_name, levels = coef_pcnt$Var_name[order(coef_pcnt$val)])
  
  p1<-ggplot(coef_pcnt)
  p1<- p1 + geom_hline(yintercept = 0, colour=gray(1/2), lty=2)
  p1<- p1 + geom_linerange(aes(x=Var_name, ymin=lowCI, ymax=highCI), lwd=1.5, position = position_dodge(width=1/2))
  p1<- p1 + geom_pointrange(aes(x= Var_name, y=coef_av, ymin=lowCI, ymax=highCI), lwd=1, position=position_dodge(width=1/2), shape=21, fill="White")
  p1<- p1 + scale_y_continuous(breaks=seq(-8, 4, 2)) +theme_bw() + labs(y = "Annual Population Change (%)", x = "Variable") + theme(legend.title=element_blank(), text = element_text(size=20),axis.title.x = element_text(margin = unit(c(5, 5, 0, 0), "mm")))
  p1<-p1 + coord_flip()
  print(p1)
  
  
#coef_pcnt$Class<-"Birds"  
#coef_pcnt$Class<-"Mammals"  

#coef_pcntb<-coef_pcnt
#coef_pcntm<-coef_pcnt  


```

```{r, eval=FALSE, echo=FALSE}
coef_both<-rbind(coef_pcntb[,c(1,2,3,4,7)], coef_pcntm[,c(1,2,3,4,7)])

coef_both$Var_name
coef_both$Var_name[coef_both$Var_name == "(Intercept)"] <- "Intercept"
coef_both$Var_name[coef_both$Var_name == "mean_slope_scale"] <- "MTC"
coef_both$Var_name[coef_both$Var_name == "change_rate_scale"] <- "LUC"
coef_both$Var_name[coef_both$Var_name == "change_rate_scale:mean_slope_scale"] <- "MTC*LUC"
coef_both$Var_name[coef_both$Var_name == "Bodymass_scale"] <- "BM"

#write.csv(coef_both, "Model_Average_coefs4.csv")
coefs_both<-read.csv("Model_Average_coefs4.csv")

library(ggplot2)
p1<-ggplot(coefs_both, aes(colour=Class))
p1<- p1 + geom_hline(yintercept = 0, colour=gray(1/2), lty=2)
p1<- p1 + geom_linerange(aes(x=Var_name, ymin=lowCI, ymax=highCI), lwd=2.5, position = position_dodge(width=2/3))
p1<- p1 + geom_pointrange(aes(x= Var_name, y=coef_av, ymin=lowCI, ymax=highCI), lwd=2, position=position_dodge(width=2/3), shape=21, fill="White")
p1<- p1 + scale_y_continuous(breaks=seq(-8, 6, 2), limits=(c(-9,5))) +theme_bw() + labs(y = "Annual Population Rate of Population Change (%)", x = "") + theme(legend.position="none",text=element_text(size=20),axis.text.x=element_text(size=20) , axis.title.x = element_text(margin = unit(c(5, 0, 0, 0), "mm")))
p1<- p1 + scale_color_manual(values=c("black", "black"))

print(p1)

```